Stability and accuracy of free surface time integration in viscous flows

### 2016-11-11 new_response_to_referee.tex

\begin{quotation}
  This paper deals with the (now) notorious issue of the drunken-sailor numerical instability that plagues free-surface calculations particularly where the surface deformation is the primary driver of flow and the flow rates are decaying. The same instability also occurs for stable internal boundaries where there same dynamic arguments apply.
\end{quotation}

\begin{quotation}
There have been a number of mechanisms suggested for "fixing" this instability published in recent years and this MS adds a new take to each of those (and appropriately acknowledges / contextualises them). I think this work is a significant addition to the collection of methods available and I would recommend that the MS be published. There has already been significant to-and-fro in review and this is apparent from the quality of the writing and clarity of the explanation. Nevertheless, every fresh pair of eyes brings some new suggestions and I would make the following comments (which I don't consider binding):
\end{quotation}

\begin{quotation}
I find it helpful to have a pseudo-code explanation of any implementation. There is a lot of derivation but in the end it appears that we have a relatively simple algorithm which pulls together equation (49) and equation (35), but if I am missing any subtle details, then perhaps that is indicating that a description would be helpful.
\end{quotation}

We agree that this is a useful summary of our implementation.
We have added Figure 4 which describes a pseudocode implementation
of both the quasi-implicit and nonstandard finite difference scheme.

@@

\begin{quotation}
I think of the sloshing instability as a physical one in which there is a resonance between the waves on a fluid and the dynamics of the container (fluid sloshing in tankers causing capsizes, for example). So I am not sure that it is a good name for a purely numerical instability even if that is how it has been named by others.
\end{quotation}

This point is well taken. We now note that this instability has been called the ``sloshing'' instability,
but in the bulk of the text we refer to it as the ``drunken sailor'' instability.

\begin{quotation}
Another example I can think of where this stabilization might be useful is at an inflow boundary where the inflow condition is used to balance a global constraint (e.g. volume / mass conservation). An example of where this arises is in modeling of extensional systems where isostasy sets the lower boundary to have a constant (average) pressure through time. There are plenty of spurious modes that can arise in the inflow condition but one instability that is often important is similar to the drunken-sailor and can be supressed by small timesteps.
\end{quotation}
This sounds like an interesting application.
Indeed, there are several situations that exhibit similar instabilities, 
and may be good targets for similar stabilization methods 
(including fluid-structure-interactions and elastoplastic deformation).
That being said, we have not analyzed those applications in this context,
and are unprepared to discuss specific commonalities.
Thus we feel this is outside of the scope of this manuscript.

@@

\begin{quotation}
Following on, we sometimes fix those sorts of problems by under-relaxing the boundary deformation over time with very little mathematical justification for doing so. I would be interested to know if this is a poor-man's equivalent to the solution you propose.
\end{quotation}

We would suggest that under-relaxing boundary deformation is indeed a poor-man's equivalent to our NSFD scheme,
and would love to hear more about this practice.
However, we are not aware of discussion of this in the literature or open-source codebases, and so can't comment further or address it in the manuscript.

### 2016-11-13 new_response_to_referee.tex

\begin{quotation}
Following on, we sometimes fix those sorts of problems by under-relaxing the boundary deformation over time with very little mathematical justification for doing so. I would be interested to know if this is a poor-man's equivalent to the solution you propose.
\end{quotation}

We would suggest that under-relaxing boundary deformation is indeed a poor-man's equivalent to our NSFD scheme,
and would love to hear more about this practice.
However, we are not aware of discussion of this in the literature or open-source codebases, and so can't comment further or address it in the manuscript.

\begin{quotation}
   Your method remains formally 1st order. Are higher-order discretisations of the mesh deformation helpful in this case or do they suffer the same / worse problems ?
\end{quotation}

It is true that our method is a first-order method in time.
Our analysis of the instability is purely for the timestepping,
so higher-order spatial discretizations should suffer from the identical instability.
We have added a note to this effect in the manuscript.
